---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r loaddp, echo=TRUE,results='hide'}
library('dplyr',quietly=TRUE, verbose=FALSE,warn.conflicts=FALSE)
```

```{r}
if(!file.exists("./activity.csv")) {
    unzip("./activity.zip")
}
movedata <- read.csv('./activity.csv')
```



## What is mean total number of steps taken per day?
```{r}
by_day <- group_by(movedata,date)
day_totals <- summarise(by_day, total_steps=sum(steps,na.rm=TRUE))
```

Histogram - Steps Per Day

```{r echo=TRUE}
hist(day_totals$total_steps,freq=7,breaks=11, xlab='Total Steps Per Day',ylab='Count',main='Histogram - Total Steps per Day')
```

Mean Total Steps per Day
``` {r echo=TRUE}
mean(day_totals$total_steps)
```

Median Total Steps Per Day
``` {r echo=TRUE}
median(day_totals$total_steps)
```



## What is the average daily activity pattern?

```{r echo=TRUE}
by_int <- group_by(movedata,interval)
interval_avg = summarize(by_int, mean_steps=mean(steps, na.rm=TRUE))
plot(interval_avg,type='l',xlab='interval',ylab='average steps',main='Average Daily Activity Pattern')
```

The 5-minute interval, on average across all days in the dataset, with the max number of steps:
```{r echo=TRUE}
filter(interval_avg, mean_steps == max(interval_avg$mean_steps))$interval
```


## Imputing missing values

Data set rows with missing values:

```{r echo=TRUE}
length(subset(movedata,is.na(movedata$steps))$steps)
```

Replace missing values with the average value for the interval, computed from the non-missing values.

```{r echo=TRUE}
md2 <- movedata
for(i in 1:dim(md2)[1]) {
  if(is.na(md2[i,1])) {
    imputed <- subset(interval_avg, interval == md2[i,3])[2]
    md2[i,1] <- imputed
  }
}
```

Histogram - Steps Per Day with Imputed Values

```{r echo=TRUE}
by_day <- group_by(md2,date)
day_totals <- summarise(by_day, total_steps=sum(steps,na.rm=TRUE))

hist(day_totals$total_steps,freq=7,breaks=11, xlab='Total Steps Per Day',ylab='Count',main='Histogram - Total Steps per Day')
```

Mean Total Steps per Day with Imputed Values
``` {r echo=TRUE}
mean(day_totals$total_steps)
```

Median Total Steps Per Day with Imputed Values
``` {r echo=TRUE}
median(day_totals$total_steps)
```

## Are there differences in activity patterns between weekdays and weekends?
